---
title: "NIRtools"
author: "Ricardo de Oliveira Perdiz"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NIRtools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE, cache = TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Overview

NIRTools was specifically designed to help plant taxonomist and ecologists in documenting and building near infrared (NIR) spectroscopy subsets from a dataset based on parameter files that will serve as a metadata of each subset.


# What is NIR?

Read this [paper](http://www.scielo.br/scielo.php?script=sci_arttext&pid=S0103-50532003000200006).

# Installation

You can install the development version from GitHub with the package `devtools`

```
install.packages("devtools")
library(devtools)
devtools::install_github("ricoperdiz/NIRtools", build_vignettes = TRUE)
```

# Quick start

Let's take a look at our example dataset and use it to build our first subset.

First, load the package and check the first rows of it.

```{r}
library(NIRtools)
head(nir_data)[,1:10]
```

We have a column `especimenid` that specifies the sample id number, a columm `SP1` that specifies our sample identity (in this case, it is the number of the species, a hypothetical number); a column `read` that specifies the leaf/leaflet id, that is, for each leaf/leaflet, we have a different value; a column `face` that specifies which surface was used to obtain the data; and at least, a group of columns that starts with letter _X_ and that refer to NIR spectra.

There are samples that have more than one read per surface. So, suppose we want to create a subset that contains only read means for both surfaces and containing all NIR spectra variables, that is, for each specimen, we would calculate mean for all reads per surface per NIR spectra variable. Our first step would be to create a parameter file that will contain all specifications of it. Let's use the function `write_NIRparams`.

In this example, my first subset will be called `test`, and  I will use mean of reads, only abaxial surface, and I will inform which column names refer to surface(`face`), individual (`especimenid`), and species/group (`SP1`).

```{r, eval = FALSE}
write_NIRparams(
  file = 'test',
  wd = getwd(),
  reads = 'mean',
  surface = 'abaxial',
  nir_variables = 'all',
  surface_id = 'face',
  individual_id = 'especimenid',
  group_id = 'SP1',
  nir_id = 'X')
```

After executing the function, a message will appear confirming the action and telling where the file has been saved.

Let's take a look at this file.

```{r}
readLines('test-NIR.txt')
```

After having our parameter file specification done, you can use function `read_NIRparams` to have your parameter file read as a dataframe.

```{r}
read_NIRparams('test-NIR.txt')
```

At last, let's use the parameter file to build a subset based on our data file. For that, let's use function `build_NIRdataset`

```{r, cache = TRUE}
subset01 <- build_NIRdataset(df = nir_data, params_file = 'test-NIR.txt', wd = '.')
```

Note that you can also choose to save a RDS and/or a txt file with this subset, you just need to choose `TRUE` instead of the default option `FALSE`.

Let's check our subset01.

```{r}
head(subset01)[,1:10]
```

# Tutorial

After our brief quick start, let's take a deeper look at NIRtools functionality. 

Suppose you want to build many subsets at once, with different combinations of values for reads and surfaces. 

In this example, I will make use of a few packages of [tidyverse](https://www.tidyverse.org/). I suggest you to learn how to use these packages because they are easy to manipulate and make reading R code easier.

## Using parameter file

First, let's create a vector of subset names.

```{r, cache = TRUE}
library(dplyr)
library(magrittr)
library(purrr)

subset_names <- sprintf('subset%02d', 1:3)
surface_values <- c('abaxial', 'adaxial', 'both')
surface_id <- 'face'
reads_values <- c('all', 'all', 'mean')
individual_id <- 'especimenid'
group_id <- 'SP1'
nir_id <- 'X'

metadata <- data_frame(subset_names, surface_values, reads_values, individual_id, surface_id, group_id, nir_id)
```

Now, let's create several parameter files at once, using all columns from our object `metadata`. Each one of its columns, as you can see above, contain all necessary data to create our parameter files.

```{r, eval = FALSE}
pwalk(list(subset_names, surface_values, reads_values, individual_id, surface_id, group_id, nir_id), ~write_NIRparams(file = ..1, wd = '.', surface = ..2, reads = ..3, individual_id = ..4, surface_id = ..5, group_id = ..6, nir_id = ..7))
```

The function `purrr::pwalk` takes a list of arguments and uses them as input for function ~NIRtools::write_NIRparams` whose result is having our three parameter files saved in our current working directory.

### Building subsets one by one

```{r}
subset_files <- list.files(pattern = 'subset')
# subset 01
subset01 <- build_NIRdataset(nir_data, subset_files[1])
# subset 02
subset02 <- build_NIRdataset(nir_data, subset_files[2])
# subset 03
subset03 <- build_NIRdataset(nir_data, subset_files[3])
```

Let's have a look at them.

```{r}
dim(subset01)
```

```{r}
dim(subset02)
```

```{r}
dim(subset03)
```

What if I do not remember which dataset refers to what? You have your parameter files to check. For example, what does subset03 mean?

```{r}
readLines('subset03-NIR.txt')
```

Or let's have it as a dataframe.

```{r}
read_NIRparams('subset03-NIR.txt')
```

### Building subsets all at once

Now, let's use the parameter files to build all our subsets at once.

```{r}
subsets <- walk2(list(nir_data), subset_files, ~build_NIRdataset(df = .x, .y))
subsets %>% str
```

The result of using function `walk2` is a list of dataframes, each one with the desired subset.

Another option, and easier to deal with, in my opinion, would be to have our parameter files stored as a dataframe, and then create a column in it that would contain our subsets. Let's take a look below to see how it is done.

```{r}
nir_params <- map_dfr(subset_files, read_NIRparams)
```

At last, let's have some magic. Let's create a column called `data` in `nir_params` and inside this column, we will store our subset in it. This is called a nested dataframe.

```{r}
nested_df <-
  nir_params %>% 
  mutate(
    main_dataset = list(nir_data),
    data = map2(main_dataset, subset_files, ~build_NIRdataset(df = .x, .y))
  )
```

Let's check our nested dataframe.

```{r}
# subset 01
dim(nested_df$data[[1]])
```

```{r}
# subset 02
dim(nested_df$data[[2]])
```

```{r}
# subset 03
dim(nested_df$data[[3]])
```

Notice that the values are exacly as the ones shown by the objects `subset01`, `subset02`, and `subset03`.

## Using a dataframe containing parameter specification

In construction.